Growing libraries of concepts with wake-sleep program induction

### 2018-07-20 talk.tex
PDF

\begin{frame}{Wake phase}
  \centering\begin{tikzpicture}[scale=1.3]
    \node at (0,0) (d){DSL};
    \node at ([yshift = -2cm]d) (t){$\text{Task}$};

    \node at ([xshift = 2cm]t) (nn){
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);

    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};

    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

    \node[align=left] at (7,-1) (f) {Frontier\\{\small (set of programs)}};
    \draw [->  ] (s.east) -- (f.west);

    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};

    \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};

  \end{tikzpicture}

\begin{frame}{Sleep-G Phase}
  Fragment grammar (\textbf{O'Donnell 2015})
  \vspace{1cm}

  \centering\begin{tikzpicture}
    \node at (0,0) (f1){Frontier$_1$};
    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
    \node at ([yshift = -0.7cm]f2.south) (ff){\textbf{$\vdots$}};
    \node at ([yshift = -1.2cm]f2.south) (ff){\textbf{$\vdots$}};
    \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};

    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
    \node[anchor=north] at (c.north) {Compression};

    \draw [-> ] (f1.east) -- (c.west|-f1.east);
    \draw [-> ] (f2.east) -- (c.west|-f2.east);
    \draw [-> ] (f3.east) -- (c.west|-f3.east);

    \node[right](d) at ([xshift = 1.2cm,yshift = 0.7cm]c.east) {DSL $\mathcal{D}$};
    \node[right](t) at ([xshift = 1.2cm,yshift = -0.7cm]c.east) {Weights $\theta$};
    \draw [-> ] (c.east) -- (d.west);
    \draw [-> ] (c.east) -- (t.west);

    \node at (c.center) {
\begin{tikzpicture}[scale=0.7]
    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
    %%     \footnotesize Program:\\
    %%     \code{($\lambda$ (x) (+ (- x) 1))}
    %% \end{tabular}};
    %\node at (,0.5) {\code{cons}};
%    \node [rotate=90] at (-2.3,-0.5) {\small program};

          \node(l1) at (0,0) {};
  \node[color=pop3](p1) at (-1,-1) {\code{+}};
  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
  \node(x1) at (0,-1) {\code{1}};
  \draw[color=pop3] (l1.south) -- (p1.north);
  \draw[color=pop3] (l1.south) -- (n1.north);
  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);

  \node(t) at (-0.5,0.5) {};
  \draw (l1.south) -- (t.south);
  \node(c) at (-1.5,-0.2) {\code{cons}};
  \draw (t.south) -- (c.north);

%    \draw  (l1.south) -- (-0.5,0.5);

  %% \node(c) at (-0.5,-1.5) {\code{-}};
  %% \node(z) at (0.5,-1.5) {\code{x}};

  %% \draw (0,-1) -- (c.north);
  %% \draw (0,-1) -- (z.north);

  \begin{scope}[shift={(-1,-2.5)}]
      \node(l1) at (0,0) {};
  \node[color=pop3](p1) at (-1,-1) {\code{+}};
  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
  %\node(x1) at (0,-1) {};
  \draw[color=pop3] (l1.south) -- (p1.north);
  \draw[color=pop3] (l1.south) -- (n1.north);
  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);

  \node(c) at (-0.5,-1.5) {\code{car}};
  \node(z) at (0.5,-1.5) {\code{z}};

  \draw (0,-1) -- (c.north);
  \draw (0,-1) -- (z.north);

%  \node [rotate=90] at (-2.3,-0.7) {\small program};

  \end{scope}

\begin{scope}[shift={(0,-5)}]
  \node[pop3](p1) at (-1,-1) {\code{+}};
  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
  \node[pop3](a) at (0,-1) {\code{ }};
  %\node(x1) at (0,-1) {};
  \draw[pop3] (0,0) -- (p1.north);
  \draw[pop3] (0,0) -- (n1.north);
  \draw[pop3] (-0.55,-0.4) -- (a.north);
%  \node [rotate=90] at (-2.3,-0.7) {\small fragment};

  \end{scope}

\end{tikzpicture}
    };

    \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};

    \end{tikzpicture}

\begin{frame}{Sleep-R Phase}
Train recognition model on two sources of data:
  \begin{itemize}
  \item Samples from generative model
    \item Programs found during waking
    \end{itemize}

\begin{frame}{Sleep-R: Train on samples from generative model}
\centering  \begin{tikzpicture}[scale=1.2]
    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
    \node at ([xshift = 3cm]d.east) (p2){program};
    \node at ([yshift = 1.5cm]p2) (p1){program};
    \node at ([yshift = -1.5cm]p2) (p3){program};

    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
    \draw[squiggle,-> ] (d.east) -- (p1.west);
    \draw[squiggle,-> ] (d.east) -- (p3.west);

    \node at ([xshift = 2cm]p1.east) (t1){task};
    \node at ([xshift = 2cm]p2.east) (t2){task};
    \node at ([xshift = 2cm]p3.east) (t3){task};
    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);

    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
  \end{tikzpicture}

\begin{frame}{Sleep-R: Train on programs found during waking}
\centering  \begin{tikzpicture}[scale=1.2]
    \node at (0,0) (f1){Frontier$_1$};
    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};

    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};

    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);

    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};

    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
  \end{tikzpicture}

@@

\begin{frame}{Text editing}
  In the style of FlashFill (Gulwani 2012)

  \centering  \includegraphics[width = 5cm]{textColumn.png}

  SyGuS problems: solves 3\% before learning, vs 74\% after learning. Best prior work: 80\%

@@

\begin{frame}{Symbolic regression from visual input}
\centering\includegraphics[width = 5cm]{symbolicRegression.png}
\end{frame}

@@

\begin{frame}{Conclusion}

  \textbf{Takeaways:}
  \begin{itemize}
  \item Language of thought models can be engineered with fewer prior assumptions and less hand engineering
    \item Humans can flexibly adapt to new problem domains --- DreamCoder takes a step in this direction
  \end{itemize}
  \pause

@@

  \begin{itemize}
  \item A real solution to the problem of searching in the language of thought: recognition model can only go so far
    \pause
    \item Generative modeling via probabilistic programs: input/output mappings do not capture many of our concepts
    \end{itemize}

>>>>

@@

\begin{frame}{List functions: Learning curves on hold out tasks}

  \begin{center}
      \includegraphics[width = 6.5cm]{figures/listLearningCurve.eps} 
    \end{center}

Learning curves for DreamCoder both with (\orange{in orange}) and without
    (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved w/ 10m timeout. Dashed lines: Average solve time, averaged only over tasks that are solved.

@@

%% \begin{frame}{Text editing}
%%   In the style of FlashFill (Gulwani 2012)

%%   \centering  \includegraphics[width = 5cm]{textColumn.png}

%%   SyGuS problems: solves 3\% before learning, vs 74\% after learning. Best prior work: 80\%

%%   \end{frame}

%% \begin{frame}{Symbolic regression from visual input}
%% \centering\includegraphics[width = 5cm]{symbolicRegression.png}
%% \end{frame}

@@

\begin{frame}{Contributions}

  \textbf{Takeaway:}
  \begin{itemize}
    \item Humans flexibly adapt to diverse sets of new problem domains --- DreamCoder takes a step in this direction
  \end{itemize}
  \pause

@@

  \begin{itemize}
  \item More human-like learning: intelligently composing new tasks, 
    \pause
  \item Theory learning
    \item Planning
    \end{itemize}
